Deriving Multi-Headed Projective Dependency Parses \\ from Link Grammar Parses

### 2014-06-03 doc/LinksToDAG.tex
small merge with the paper. working on the conll_analysis.py scripts.
Added a table to the appendix section. The English Link Types.
Changed the table in the appendix to use coarse link labels. Made the table smaller.
updated the paper

As recently discussed by \newcite{gomezrodriguez-nivre-2013}, one might impose various requirements on the parse graph:
\begin{itemize}
\item {\sc Single-Head}: each word has $\leq 1$ parent
\item {\sc Acyclic}: there are no directed cycles
\item {\sc Connected}: each pair of words has a undirected path between them
\item {\sc Reachable}: each word can be reached from 0 by a directed path
  \noteJE{note that this implies connected?  or does connected not allow 0 words on the path, i.e., does it restrict 0 to one child, and if not, is there a principle for that?  If no one, mention in footnote 1.}
\item {\sc Planar}: edges may not ``cross'' \\ (if there are edges between $u,v$ and between $i,j$, where $u < i < v$, then $u \leq j \leq v$)
\end{itemize}
It is common to impose all of these requirements at once, leading to a {\em projective dependency parser} that produces projective trees rooted at 0.\footnote{{\sc Projective} basically means {\sc Planar} + {\sc Reachable}.}  However, parsing algorithms can be devised that relax any or all of the requirements \cite{gomezrodriguez-nivre-2013}.

@@

\section{Data Sets}
The data used for this work came from English bnews CoNLL dependency training data \noteJH{cite?} \noteJE{wait a minute, the bnews is not from CoNLL!  It is merely in CoNLL format.  It is from Ariya Rastrow, and is automatically parsed, I think.  We should be comparing on hand-parsed data.}  and the Russian National Corpus\footnote{\url{http://ruscorpora.ru/en/}} \noteJE{give citation}.

We generated link parses using the AbiWord/CMU link grammar parser version 5.0.8 \cite{LINKPARSER-2014}.  The parser's coverage is not perfect: we obtained complete, connected parses for only !!! (of !!!) English sentences and !!! (of !!!) Russian sentences, discarding the other sentences. \noteJE{fill these !!! in automatically} These two languages have the most complete lexicons at present, although lexicons for !!! other languages are available.

@@

\subsection{Discarding Incomplete Link Parses}

\NoteJE{ideally we could cut this section given what I wrote above.  However, you seem to be saying that we did train on sentences with dropped words as long as the parses were connected.  Didn't we decide to drop those sentences too?}

Not all of the data was used for this work. From the given sentences, we discarded those that the link parser could not process and returned a parse with nodes disconnected from the root. \input{figure/conll_analysis_sentences_disconnected} of the English sentences were removed in this way. The ILP was run only on the connected parses to produce a corpus of directionalized link parses.

For our analysis of these parses, we further discarded those that had dropped words, where the link parser could not attach links to every word in the sentence. About \input{figure/conll_analysis_sentences_dropped}of the non-disconnected sentences were removed, leaving us working only with complete link parses.

\begin{figure}[ht!]
  \centering
  \small
  \input{figure/conll_analysis_sentences.tex}
  \caption{\small The English sentences used for this work.}
\end{figure}

\noteJH{The stability results come from the second row of this table. The analysis later in the paper uses the third row. Putting this section here might be confusing to readers?}

@@

\section{Experiments and Results}

We solved our ILP problem using the SCIP Optimization Suite \cite{achterberg2009scip}, encoding it using the ZIMPL language \cite{Koch2004}.  
\NoteJE{give runtime here}

\subsection{Stability of Results}
\begin{figure}[ht!]
  \small
  \centering
  \includegraphics[width=\linewidth, keepaspectratio=true]{figure/precision_recall.png}
  \caption{\small Convergence to the results obtained on the largest run.  Precision relative to that run is very high even for small training sizes.  Recall grows as more link types are observed.}
\end{figure}
We report on the convergence of the results of our ILP with increasing data. With increasing trials of sentences, we measure how similar the subsequent directional assignments would be with each other. Taking the largest run as the answer key, we compared how much the smaller runs deviated from it. 
We recorded the precision of whether the assignments in the smaller runs could be found in those of the largest run, and similarly the recall of whether the assignments in the largest run could be found in the smaller runs. \noteJE{this is not clear.  are we comparing the direction table or the parses?  If the parses, which parses?}

@@

\begin{figure*}[ht!]
  \centering
  \input{figure/sentences.tikz}
  \caption{\small Example Parses: The top blue arcs represent conll dependencies; the bottom red arcs represent directionalized link grammar parses. See the Appendix for more details.}
  \label{fig:parses}
\end{figure*}

@@

Of the sentences with complete parses, about \input{figure/conll_analysis_sentences_multiheaded}
\noteJE{why ``about''?}
had at least one multiheaded word.

@@

>>>>

As recently discussed by \newcite{gomezrodriguez-nivre-2013}, one might impose various requirements on the parse graph:
\begin{itemize}

\item {\sc Single-Head}: each word has $\leq 1$ parent
\item {\sc Single-Head}: each word has at most one parent
\item {\sc Acyclic}: there are no directed cycles
\item {\sc Connected}: each pair of words has a undirected path between them
\item {\sc Reachable}: each word can be reached from 0 by a directed path
  \noteJE{note that this implies connected?  or does connected not allow 0 words on the path, i.e., does it restrict 0 to one child, and if not, is there a principle for that?  If no one, mention in footnote 1.}
\item {\sc Planar}: edges may not ``cross'' \\ (if there are edges between $u,v$ and between $i,j$, where $u < i < v$, then $u \leq j \leq v$)
\end{itemize}
It is common to impose all of these requirements at once, leading to a {\em projective dependency parser} that produces projective trees rooted at 0.\footnote{{\sc Projective} basically means {\sc Planar} + {\sc Reachable}.}  However, parsing algorithms can be devised that relax any or all of the requirements \cite{gomezrodriguez-nivre-2013}.

@@

\section{Data Sets}
The data used for this work came from English Penn Tree Bank CoNLL dependency training data \noteJH{cite?} 
%\noteJE{wait a minute, the bnews is not from CoNLL!  It is merely in CoNLL format.  It is from Ariya Rastrow, and is automatically parsed, I think.  We should be comparing on hand-parsed data.}  
and the Russian National Corpus\footnote{\url{http://ruscorpora.ru/en/}} \cite{russiancorpus} \noteJE{give citation}.

We generated link parses using the AbiWord/CMU link grammar parser version 5.0.8 \cite{LINKPARSER-2014}.  The parser's coverage is not perfect: we obtained complete, connected parses for only \input{figure/conll_analysis_sentences_total.tex}of (\input{figure/conll_analysis_sentences_original.tex}) English sentences and only 206 (of 10,000) Russian sentences, discarding the other sentences. 
%\noteJE{fill these !!! in automatically} 
These two languages have the most complete lexicons at present, although lexicons for seven other languages are available.

@@

%\subsection{Discarding Incomplete Link Parses}

%\NoteJE{ideally we could cut this section given what I wrote above.  However, you seem to be saying that we did train on sentences with dropped words as long as the parses were connected.  Didn't we decide to drop those sentences too?}

%Not all of the data was used for this work. From the given sentences, we discarded those that the link parser could not process and returned a parse with nodes disconnected from the root. \input{figure/conll_analysis_sentences_disconnected} of the English sentences were removed in this way. The ILP was run only on the connected parses to produce a corpus of directionalized link parses.

%For our analysis of these parses, we further discarded those that had dropped words, where the link parser could not attach links to every word in the sentence. About \input{figure/conll_analysis_sentences_dropped}of the non-disconnected sentences were removed, leaving us working only with complete link parses.

%\begin{figure}[ht!]
%  \centering
%  \small
%  \input{figure/conll_analysis_sentences.tex}
%  \caption{\small The English sentences used for this work.}
%\end{figure}

%\noteJH{The stability results come from the second row of this table. The analysis later in the paper uses the third row. Putting this section here might be confusing to readers?}

@@

\section{Experiments and Results}
\begin{figure*}[ht!]
  \centering
  \input{figure/sentences.tikz}
  \caption{\small Example Parses: The top blue arcs represent conll dependencies; the bottom red arcs represent directionalized link grammar parses. See the Appendix for more details.}
  \label{fig:parses}
\end{figure*}

We solved our ILP problem using the SCIP Optimization Suite \cite{achterberg2009scip}, encoding it using the ZIMPL language \cite{Koch2004}.  
\NoteJE{give runtime here}\NoteJH{A run of 8,192 sentences takes 1743 seconds}

\subsection{Stability of Results}
\begin{figure}[ht!]
  \small
  \centering
  \includegraphics[width=\linewidth, keepaspectratio=true]{figure/precision_recall.png}
  \caption{\small Convergence to the results obtained on the largest run.  Precision relative to that run is very high even for small training sizes.  Recall grows as more link types are observed.}
\end{figure}
We report on the convergence of the results of our ILP with increasing data. With increasing trials of sentences, we measure how similar the subsequent directional assignments would be with each other. Taking the direction table of the largest run as the answer key, we compared how much the solutions of the smaller runs deviated from it. 
We recorded the precision of whether the assignments in the smaller runs could be found in those of the largest run, and similarly the recall of whether the assignments in the largest run could be found in the smaller runs. 
%\noteJE{this is not clear.  are we comparing the direction table or the parses?  If the parses, which parses?}

@@

@@

Of the English sentences with complete parses, \input{figure/conll_analysis_sentences_multiheaded}had at least one multiheaded word.

@@

\input{figure/link_analysis_coarse_table.tex}
